---
title: "Project 1"
output: html_document
author: Joseph Cieselwitz, Garrett Carr, Yuko Ketch, Luke McDowell
---

```{r setup, include=FALSE}
library(tidyverse)
library(car)
library(ggfortify)

```


## Background & Introduction
The World Health Organization’s purpose for creating a new life-expectancy dataset was to add demographic variables, income composition and mortality rate as factors affecting life expectancy. We saw this as an opportunity to explore the effect education has had on life expectancy. Our data set was sourced from Kaggle using the World Health Organization’s data. 

The Covid-19 pandemic has highlighted the importance of supporting public health in the developing world. We believe this dataset can help inform our understanding of the relationship between education and greater health, which would potentially aid in worldwide efforts for greater physical well-being. 


```{r cars}
WHO <- read_csv("Life Expectancy Data.csv", col_names = TRUE) %>% 
  rename(Life_Exp = `Life expectancy`)
WHO
WHO_2015 <- WHO %>% filter(Year == 2015) %>% 
  select(Country, Life_Exp, Schooling) %>%
  filter(!(is.na(Schooling)))
WHO_2015
```

## Including Plots

Our null and alternative hypotheses for this study will be the following:
	Ho = No correlation between years of schooling and life expectancy
	Ha = Strong, positive, linear correlation between years of schooling, and life expectancy


We plan to use a simple linear regression model with the R programming language to carry out our computations. We will use known regression diagnostic tools to assess assumptions, and carry out transformations as necessary. Our team will meet in person, and over video calls to complete our study. 


```{r pressure}
# Scatterplot
life_base <- ggplot(data=WHO_2015, aes(x = Schooling, y = Life_Exp)) + 
  geom_point()

WHO_2015
summary(WHO_2015)

ggplot(data=WHO_2015, aes(x = Life_Exp)) +
  geom_histogram(aes(y=..density..), binwidth = 4)


```


```{r Assumptions}
# Linearity
life_base

life_lm <- lm(data = WHO_2015, Life_Exp ~ Schooling)
summary(life_lm)
WHO_2015$residuals <- life_lm$residuals

RvF <- autoplot(life_lm, which = 1, nrow = 1, ncol = 1) + 
  theme(aspect.ratio=1)
RvF

```


```{r Assumptions}
# Boxplot of residuals
ggplot(data = WHO_2015, aes(y=residuals)) + geom_boxplot()

# Histogram of residuals
stop_hist <- ggplot(data = WHO_2015, aes(x=residuals)) +
  geom_histogram(aes(y=..density..), binwidth = 3) +
  #Overlay the normal curve with this mean and sd
  stat_function(fun = dnorm,
                size = 2,
                args = list(mean = mean(WHO_2015$residuals), 
                            sd = sd(WHO_2015$residuals)))

stop_hist

# QQ Plot
stop_QQ <- autoplot(life_lm, which = 2, ncol = 1, nrow = 1)
stop_QQ

shapiro.test(life_lm$residuals)

```


```{r Tranformation}
bc <- boxCox(WHO_2015$Life_Exp ~ WHO_2015$Schooling)
bc$x[which.max(bc$y)]

WHO_2015$Life_Sq <- WHO_2015$Life_Exp^2
life_lm_sq <- lm(data = WHO_2015, Life_Sq ~ Schooling)
summary(life_lm_sq)
WHO_2015$Residuals_Sq <- life_lm_sq$residuals


```

